---
title: "Task 2: How Kenyan maternal/child indicators (stunting, skilled birth attendance, immunisation) vary by region and wealth quintile."
author: "Kevinson Mwangi & Elphas Abok"
date: today
format: html
---

## Loading the KDHS Data 2022 (KR Record)
```{r}
#| label: package-setup
#| message: false
#| warning: false
#| cache: false
#| include: true

rm(list = ls())

# Packages
pacman::p_load(
  tidyverse, haven, labelled, epiDisplay, survey,
  gtsummary, extrafont, flextable, rKenyaCensus, sf, rineq)

# Data
df <- read_dta("../Mini project/Data/KEKR8BDT/KEKR8BFL.DTA")

```


```{r}
#| label: plotting-functions-pointrange
#| message: false
#| warning: false
#| cache: false
#| include: true

plot_pointrange <- 
    function(data, x_var, estimate, lower, upper, 
    reference_line = NULL, y_limits = c(5, 50),
    y_breaks = seq(5, 50, 5),
    y_label = "Weighted prevalence (%)") {

  x_var <- rlang::ensym(x_var)
  estimate <- rlang::ensym(estimate)
  lower <- rlang::ensym(lower)
  upper <- rlang::ensym(upper)

  p <- data |>
    ggplot(
      aes(
        x = forcats::fct_reorder(!!x_var, !!estimate, .desc = TRUE),
        y = !!estimate)) +
    geom_pointrange(
      aes(
        ymin = !!lower,
        ymax = !!upper),
      size = .3,
      color = "gray30") +
    scale_y_continuous(
      limits = y_limits,
      breaks = y_breaks) +
    labs(
      x = NULL,
      y = y_label) +
    theme_bw(
      base_family = "Candara",
      base_size = 14) +
    theme(
      axis.text = element_text(color = "black"),
      axis.text.x = element_text(angle = 90, vjust = .5))

  if (!is.null(reference_line)) {
    p <- p +
      geom_hline(
        yintercept = reference_line,
        lty = 2,
        color = "blue")
  }

  return(p)
}

```


```{r}
#| label: plotting-functions-maps
#| message: false
#| warning: false
#| cache: false
#| include: true

plot_map <- function(
    data, fill, breaks, labels, legend_title, subtitle, palette = "YlGnBu", direction = 1) {

  fill <- rlang::ensym(fill)

  data |>
    mutate(
      fill_cat = cut(
        !!fill,
        breaks = breaks,
        labels = labels,
        right = FALSE,
        include.lowest = TRUE)) |>
    ggplot() +
    geom_sf(
      aes(fill = fill_cat),
      color = "gray30") +
    scale_fill_brewer(
      palette = palette,
      direction = direction) +
    labs(
      fill = legend_title,
      subtitle = subtitle) +
    theme_minimal(base_family = "Candara", base_size = 14) +
    theme(
      axis.text = element_blank(),
      plot.subtitle = element_text(hjust = .6, vjust = -8),
      legend.background = element_rect(color = "gray70"),
      legend.title = element_text(size = 10.5),
      legend.text = element_text(size = 10.5),
      legend.position = c(0.15, 0.16),
      panel.grid = element_blank()
    )
}

```


## Data preparation
```{r}
#| label: data-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

## Restrict the analysis to the youngest alive child, living with the responded 
df <- df |>  
    mutate(
        age_months = v008 - b3) |> 
    filter(
        b5 == 1, # child is alive
        b9 == 0, # child lives with the respondent
        !is.na(v003) # mother's line number must be available
        ) |>
    group_by(
        v001, 
        v002, 
        v003 # grouping by cluster, household, and mothers line number
    ) |> 
    mutate(
        bidxr = 
            rank(bidx, ties.method = "first")) |> 
    filter(
        bidxr == 1 # Keeping the youngest child 
        ) |> 
    ungroup()

df <- df |> 
    mutate(
        wt = v005/1000000, # sample weight
        wealth_cont = v191/100000,
        wealth_q = v190
    )
```


The datasets includes a total of `r dim(df)[[1]]` children from `r length(unique(df$v001))` clusters across Kenya. 

## The ouctomes of interest
### Stunting
#### National weighted n and prevalence of stunting

```{r}
#| label: fig-stunting
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Histogram of HAZ scores"

df_stunting <- df |> 
    filter(
        age_months > 5 & age_months < 60) |> 
  mutate(
    hw70 = ifelse( #hw70 is the height-for-age zscore and should be divided by 100
        hw70 >
         1000, NA, hw70),
    haz = hw70/100,
    
    stunted = ifelse(haz < -2, "Stunted", "Not stunted"),
    stun = ifelse(stunted == "Stunted", 1, 0))
 
df_stunting |> 
    ggplot(
        aes(x = haz)) +
            geom_histogram(color = "white") +
            labs(
                x = "Height-for-age z-score (HAZ scores)",
                y = "Frequency") + 
    theme_bw(base_family = "Candara", base_size = 14) +
    theme(panel.grid = element_blank())
```


```{r}
#| label: survey-setting
#| message: false
#| warning: false
#| cache: false
#| include: true

# Survey setting
stunting_survey <- svydesign(
  id = ~v021, # primary sampling unit       
  strata = ~v022, # strata     
  weights = ~wt,       
  data = df_stunting,
  nest = TRUE)  

options(survey.lonely.psu = "adjust") # adjust for case where there may be 1 psu in a strata

# Weighted prevalence of stunting
tbl <- tbl_svysummary(
  stunting_survey,
  include = stunted,
  statistic = all_categorical() ~ "{n} ({p})",
  digits = all_categorical() ~ c(1, 1),
  missing = "ifany",
  missing_text = "Missing",
  label = stunted ~ "Stunting status")
```

Stunting was defined as height-for-age z-score (HAZ) < -2 based on the WHO 2006 growth reference standards. The analysis included a total of `r dim(df_stunting)[[1]]`. According to the WHO guidelines, HAZ scores range from -6 to 6 and this was ascertained for the data as shown in @fig-stunting. The data was then survey-set to account for the two-stage sampling design of DHS. The weighted n (prevalence %) of stunting among the 6-59 months old children was `r tbl$table_body[3, 6][[1]]`.


#### Does stunting vary by region (subnational levels/counties)?

```{r}
#| label: fig-prev-counties
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of stunting among children 6-59 months at the subnationa levels. The blue line represents the national weighted prevalence"

county_stunting <- svyby(
  ~ I(stunted == "Stunted"),
  by = ~v024,
  design = stunting_survey,
  FUN = svyciprop,
  vartype = "ci",
  na.rm = TRUE)

stunting_county <- county_stunting |> 
    mutate(
        county = v024,
        prevalence = `I(stunted == "Stunted")`,
        lb = ci_l,
        ub = ci_u) |> 
    select(county, prevalence, lb, ub)


stunting_county <- stunting_county |> 
    mutate(
        name = names(attr(county, "labels")) |> 
            str_to_title() |> 
            str_replace_all("[[:punct:]]", " "))


# Fetching Kenya counties shapefiles
kenya_shp <- rKenyaCensus::KenyaCounties_SHP |> 
    janitor::clean_names() |> 
    st_as_sf() |> 
    select(county) |> 
    mutate(
        county = str_to_title(county) |> 
            str_replace_all("[[:punct:]]", " "))

kenya_shp[kenya_shp$county == "Nairobi City", "county"] <- "Nairobi"

# Merge the prevalence data to the shp file data
merged_stunting <- kenya_shp |> 
    left_join(
        stunting_county, by = c("county" = "name")) |> 
    mutate(
        across(
            c(prevalence, lb, ub), ~round(., 3)*100))

# Plotting the prevalence across the counties
plot_pointrange(
    merged_stunting, x_var = "county", estimate = "prevalence",
    lower = lb, upper = ub,
    reference_line = 
        as.numeric(
            str_extract(tbl$table_body[3, 6][[1]],
             "(?<=\\()[0-9.]+(?=\\))")),
    y_limits = c(5, 50),
    y_breaks = seq(5, 50, 5),
    y_label = "Weighted prevalence (%)")
```


@fig-prev-counties shows the weighted prevalence of stunting among children 6-59 months at the subnational levels with the blue dotted line showing the weighted national prevalence.

```{r}
#| label: fig-prev-map
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of stunting among children 6-59 months at the subnationa levels"
        
## Prevalence maps
plot_map(
    merged_stunting, fill = prevalence,
    breaks = c(0, 10, 20, 30, 40, 50),
    labels = c("0–10", "11–20", "21–30", "31–40", "41–50"),
    legend_title = "Prevalence (%)", subtitle = NULL,
    palette = 6, direction = 1)
    
```


@fig-prev-map shows the weighted prevalence of stunting among children 6-59 months at the subnational levels. Darker shades shows increasing prevalence

#### Does stunting vary by wealth status?
##### Weighted concentration curves
```{r}
#| label: concentration-curves-data-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

cc_df <- df_stunting |> 
    filter(!is.na(stun), !is.na(wealth_cont), !is.na(wt)) |>     
    arrange(wealth_cont) |> 
    mutate(
        cum_prop = cumsum(wt) / sum(wt),
        cum_stunted = cumsum(stun*wt) / sum(stun*wt))
    

cc_df <- df_stunting |>  
    filter(!is.na(stun), !is.na(wealth_cont), !is.na(wt)) |>     
    arrange(wealth_cont) |> 
    mutate(
    pop_rank = cumsum(wt) - wt / 2, # midpoint rank within own weight mass
    pop_rank = pop_rank / sum(wt)) |> 
  group_by(wealth_cont) |>   # collapse exact ties
  mutate(pop_rank = mean(pop_rank)) |> # give tied group its average rank
  ungroup() |> 
  arrange(pop_rank) |> 
  mutate(
        cum_prop = cumsum(wt) / sum(wt),
        cum_stunted = cumsum(stun*wt) / sum(stun*wt))

cc_plot_data <- bind_rows(
  data.frame(cum_prop = 0, cum_stunted = 0),
  cc_df |> select(cum_prop, cum_stunted)) # Adding (0, 0) for plotting

```

```{r}
#| label: fig-concentration-curves-plotting
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of stunting among children 6-59 months at the subnationa levels"

## Concentration index
ci_result <- ci(
  ineqvar = df_stunting$wealth_cont,
  outcome = df_stunting$stun,
  weights = df_stunting$wt,
  type    = "CIw")

cc_plot_data |> 
  ggplot(aes(x = cum_prop, y = cum_stunted)) +
  geom_line(color = "navyblue", linewidth = 1) +
  geom_abline(
    slope = 1, intercept = 0,
     color = "#813134") +
  annotate(
    "text",
    x = 0.05, y = 0.95,
    label = paste0(
      "CI (95% CI) = ",
      round(ci_result$concentration_index, 2),
      " (",
      round(confint(ci_result)[[1]], 2), ",",
      round(confint(ci_result)[[2]], 2), ")"),
    family = "Candara",
    hjust = 0, vjust = 1,
    size = 4.0) +
  labs(
    x = "Cumulative % of children, ranked by wealth (poorest \u2192 richest)",
    y = "Cumulative % of stunted children") +
  theme_classic(base_family = "Candara", base_size = 13) +
  theme(
    panel.grid = element_blank())

```


## Immunisation

Vaccination analysis was restricted to only 12-23 and 24-35 months. Full vaccination was defined differently for these two groups of children according to the national vaccination schedule.

### Distribution of full vaccination at the national and subnational levels, children 12-23 months

Full vaccination was defined as having taken one dose of BCG vaccine,oral polio vaccine (OPV, birth dose), three doses of OPV and one dose of inactivated polio vaccine (IPV), three doses of DPT-HepB-Hib, three doses of pneumococcal conjugate vaccine (PCV), two doses of Rotavirus vaccine (RV), and One dose of measles rubella (MR) among children 12-23 months.

```{r}
#| label: immunisation-prep-1
#| message: false
#| warning: false
#| cache: false
#| include: true

# Defining fully vaccinated children 12-23 years
df_immunn_12_23 <- df |> 
    filter(
        age_months >= 12 & age_months < 24) |> 
    mutate(
        fully_vaccinated = 
            ifelse(

            (h2 %in% c(1, 2, 3) & # one dose of BCG vaccine

            h0 %in% c(1, 2, 3) & # Oral polio vaccine (birth dose)
            h4 %in% c(1, 2, 3) & # Oral polio vaccine 1
            h6 %in% c(1, 2, 3) & # Oral polio vaccine 2
            h8 %in% c(1, 2, 3) & # Oral polio vaccine 3
            h60 %in% c(1, 2, 3) & # one dose of inactivated polio vaccine

            h3 %in% c(1, 2, 3) & # Dose 1 DPT-HepB-Hib
            h5 %in% c(1, 2, 3) & # Does 2 DPT-HepB-Hib
            h7 %in% c(1, 2, 3) & # Dose 3 DPT-HepB-Hib

            h54 %in% c(1, 2, 3) & # Pneumococcal conjugate vaccine 1
            h55 %in% c(1, 2, 3) & # Pneumococcal conjugate vaccine 2
            h56 %in% c(1, 2, 3) & # Pneumococcal conjugate vaccine 3

            h57 %in% c(1, 2, 3) & # Rotavirus vaccine 1
            h58 %in% c(1, 2, 3) & # Rotavirus vaccine 2

            h9 %in% c(1, 2, 3)  # Measles rubella 1
            ), "Fully vaccinated", "Not fully vaccinated"),
        fully_vacc = ifelse(
            fully_vaccinated == "Fully vaccinated", 0, 1)
    )

```

Survey setting step

```{r}
#| label: immunisation-prep-12_23
#| message: false
#| warning: false
#| cache: false
#| include: true

immunn_survey_12_23 <- svydesign(
  id = ~v021, # primary sampling unit       
  strata = ~v022, # strata     
  weights = ~wt,       
  data = df_immunn_12_23,
  nest = TRUE)  

options(survey.lonely.psu = "adjust") # adjust for case where there may be 1 psu in a strata
```

Weighted prevalence of full vaccination - national level

```{r}
#| label: weighted-prev-12_23
#| message: false
#| warning: false
#| cache: false
#| include: true

tbl_immun_12_23 <- tbl_svysummary(
  immunn_survey_12_23,
  include = fully_vaccinated,
  statistic = all_categorical() ~ "{n} ({p})",
  digits = all_categorical() ~ c(1, 1),
  missing = "ifany",
  missing_text = "Missing",
  label = fully_vaccinated ~ "Vaccination status")

```

Weighted prevalence of full vaccination - subnational levels

```{r}
#| label: fig-prev-counties-immunisation-12_23
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of full vacciantion among children 12-23 months at the subnationa levels. The blue line represents the national weighted prevalence"

county_immunn_12_23 <- svyby(
  ~ I(fully_vaccinated == "Fully vaccinated"),
  by = ~v024,
  design = immunn_survey_12_23,
  FUN = svyciprop,
  vartype = "ci",
  na.rm = TRUE)

immunn_12_23_county <- county_immunn_12_23 |> 
    mutate(
        county = v024,
        prevalence = `I(fully_vaccinated == "Fully vaccinated")`,
        lb = ci_l,
        ub = ci_u,
        group = rep("12-23 months")) |> 
    select(county, group, prevalence, lb, ub)


immunn_12_23_county <- immunn_12_23_county |> 
    mutate(
        name = names(attr(county, "labels")) |> 
            str_to_title() |> 
            str_replace_all("[[:punct:]]", " "))


# Merge the prevalence data to the shp file data
merged_immunn_12_23 <- kenya_shp |> 
    left_join(
        immunn_12_23_county, by = c("county" = "name")) |> 
    mutate(
        across(
            c(prevalence, lb, ub), ~round(., 3)*100))

# Plotting
plot_pointrange(
    merged_immunn_12_23, x_var = "county", estimate = "prevalence",
    lower = lb, upper = ub,
    reference_line = 
        as.numeric(
            str_extract(tbl_immun_12_23$table_body[3, 6][[1]],
             "(?<=\\()[0-9.]+(?=\\))")),
    y_limits = c(0, 100),
    y_breaks = seq(0, 100, 10),
    y_label = "Weighted prevalence (%)")

```

Spatial distribution of full vacciantion among children 12-23 months at the subnationa level

```{r}
#| label: fig-prev-maps-immunn-12_23
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Spatial distribution of full vacciantion among children 12-23 months at the subnationa level"

figa <- plot_map(
  merged_immunn_12_23,
  fill = prevalence,
  breaks = c(0, 20, 40, 60, 80, 100),
  labels = c("<20", "20–39.9", "40–59.9", "60–79.9", "80–100"),
  legend_title = "Prevalence (%)",
  subtitle = "12–23 months",
  palette = 8,
  direction = -1
)

figa

```


Concentration curve for full vaccination by cumulative percentage of the population ranked by wealth index

Data preparation step

```{r}
#| label: concentration-curves-data-prep-12_23
#| message: false
#| warning: false
#| cache: false
#| include: true

cc_df_immunn_12_23 <- df_immunn_12_23 |>   
    filter(!is.na(fully_vacc), !is.na(wealth_cont), !is.na(wt)) |>     
    arrange(wealth_cont) |> 
    mutate(
    pop_rank = cumsum(wt) - wt / 2, # midpoint rank within own weight mass
    pop_rank = pop_rank / sum(wt)) |> 
  group_by(wealth_cont) |>   # collapse exact ties
  mutate(pop_rank = mean(pop_rank)) |> # give tied group its average rank
  ungroup() |> 
  arrange(pop_rank) |> 
  mutate(
        cum_prop = cumsum(wt) / sum(wt),
        cum_full_vacc = cumsum(fully_vacc*wt) / sum(fully_vacc*wt))


cc_plot_data_immunn_12_23 <- bind_rows(
  data.frame(cum_prop = 0, cum_full_vacc = 0),
  cc_df_immunn_12_23 |> select(cum_prop, cum_full_vacc)) # Adding (0, 0) for plotting

```

Generating concentration index and plotting

```{r}
#| label: fig-concentration-curves-plotting-immunn-12_23
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Concentration curves showing "

## Concentration index
ci_result_immunn_12_23 <- ci(
  ineqvar = df_immunn_12_23$wealth_cont,
  outcome = df_immunn_12_23$fully_vacc,
  weights = df_immunn_12_23$wt,
  type    = "CIw")

cc_plot_data_immunn_12_23 |> 
  ggplot(aes(x = cum_prop, y = cum_full_vacc)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_abline(
    slope = 1, intercept = 0,
     color = "#813134") +
  annotate(
    "text",
    x = 0.05, y = 0.95,
    label = paste0(
      "Concentration index (95% CI) = ",
      round(ci_result_immunn_12_23$concentration_index, 2),
      " (",
      round(confint(ci_result_immunn_12_23)[[1]], 2), ",",
      round(confint(ci_result_immunn_12_23)[[2]], 2), ")"),
    family = "Candara",
    hjust = 0, vjust = 1,
    size = 4.0) +
  labs(
    x = "Cumulative % of children, ranked by wealth (poorest \u2192 richest)",
    y = "Cumulative % of non fully vaccinated children",
    subtitle = "12-23 months") +
  theme_classic(base_family = "Candara", base_size = 13) +
  theme(
    panel.grid = element_blank())

```


### Distribution of full vaccination at the national and subnational levels, children 24-35 months

Full vaccination was defined as having taken one dose of BCG vaccine,oral polio vaccine (OPV, birth dose), three doses of OPV and one dose of inactivated polio vaccine (IPV), three doses of DPT-HepB-Hib, three doses of pneumococcal conjugate vaccine (PCV), two doses of Rotavirus vaccine (RV), and two doses of measles rubella (MR) among children 24-35 months.

```{r}
#| label: immunisation-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

# Defining fully vaccinated children 24-35 years
df_immunn_24_35 <- df |> 
    filter(
        age_months >= 24 & age_months < 36) |> 
    mutate(
        fully_vaccinated = 
            ifelse(

            (h2 %in% c(1, 2, 3) & # one dose of BCG vaccine

            h0 %in% c(1, 2, 3) & # Oral polio vaccine (birth dose)
            h4 %in% c(1, 2, 3) & # Oral polio vaccine 1
            h6 %in% c(1, 2, 3) & # Oral polio vaccine 2
            h8 %in% c(1, 2, 3) & # Oral polio vaccine 3
            h60 %in% c(1, 2, 3) & # one dose of inactivated polio vaccine

            h3 %in% c(1, 2, 3) & # Dose 1 DPT-HepB-Hib
            h5 %in% c(1, 2, 3) & # Does 2 DPT-HepB-Hib
            h7 %in% c(1, 2, 3) & # Dose 3 DPT-HepB-Hib

            h54 %in% c(1, 2, 3) & # Pneumococcal conjugate vaccine 1
            h55 %in% c(1, 2, 3) & # Pneumococcal conjugate vaccine 2
            h56 %in% c(1, 2, 3) & # Pneumococcal conjugate vaccine 3

            h57 %in% c(1, 2, 3) & # Rotavirus vaccine 1
            h58 %in% c(1, 2, 3) & # Rotavirus vaccine 2

            h9 %in% c(1, 2, 3) &  # Measles rubella 1
            h9a %in% c(1, 2, 3)   # Measles rubella 2
            ), "Fully vaccinated", "Not fully vaccinated"),
        fully_vacc = ifelse(
            fully_vaccinated == "Fully vaccinated", 0, 1)        
    )

```

Survey setting step

```{r}
#| label: immunisation-prep-24_35
#| message: false
#| warning: false
#| cache: false
#| include: true
#| 
immunn_survey_24_35 <- svydesign(
  id = ~v021, # primary sampling unit       
  strata = ~v022, # strata     
  weights = ~wt,       
  data = df_immunn_24_35,
  nest = TRUE)  

options(survey.lonely.psu = "adjust") # adjust for case where there may be 1 psu in a strata

```


Weighted prevalence of full vaccination - national level

```{r}
#| label: weighted-prev-24_35
#| message: false
#| warning: false
#| cache: false
#| include: true

tbl_immun_24_35 <- tbl_svysummary(
  immunn_survey_24_35,
  include = fully_vaccinated,
  statistic = all_categorical() ~ "{n} ({p})",
  digits = all_categorical() ~ c(1, 1),
  missing = "ifany",
  missing_text = "Missing",
  label = fully_vaccinated ~ "Vaccination status")

```

Weighted prevalence of full vaccination - subnational levels

```{r}
#| label: fig-prev-counties-immunisation-24_35
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of full vacciantion among children 24-35 months at the subnationa levels. The blue line represents the national weighted prevalence"

county_immunn_24_35 <- svyby(
  ~ I(fully_vaccinated == "Fully vaccinated"),
  by = ~v024,
  design = immunn_survey_24_35,
  FUN = svyciprop,
  vartype = "ci",
  na.rm = TRUE)

immunn_24_35_county <- county_immunn_24_35 |> 
    mutate(
        county = v024,
        prevalence = `I(fully_vaccinated == "Fully vaccinated")`,
        lb = ci_l,
        ub = ci_u,
        group = rep("24-35 months")) |> 
    select(county, group, prevalence, lb, ub)


immunn_24_35_county <- immunn_24_35_county |> 
    mutate(
        name = names(attr(county, "labels")) |> 
            str_to_title() |> 
            str_replace_all("[[:punct:]]", " "))


# Merge the prevalence data to the shp file data
merged_immunn_24_35 <- kenya_shp |> 
    left_join(
        immunn_24_35_county, by = c("county" = "name")) |> 
    mutate(
        across(
            c(prevalence, lb, ub), ~round(., 3)*100))


# Plotting the prevalence across the counties
plot_pointrange(
    merged_immunn_24_35, x_var = "county", estimate = "prevalence",
    lower = lb, upper = ub,
    reference_line = 
        as.numeric(
            str_extract(tbl_immun_24_35$table_body[3, 6][[1]],
             "(?<=\\()[0-9.]+(?=\\))")),
    y_limits = c(0, 100),
    y_breaks = seq(0, 100, 10),
    y_label = "Weighted prevalence (%)")

```

Spatial distribution of full vacciantion among children 12-23 months at the subnationa level

```{r}
#| label: fig-prev-maps-immunn-24_35
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Spatial distribution of full vacciantion among children 24-35 months at the subnationa level"

figb <- plot_map(
  merged_immunn_24_35,
  fill = prevalence,
  breaks = c(0, 20, 40, 60, 80, 100),
  labels = c("<20", "20–39.9", "40–59.9", "60–79.9", "80–100"),
  legend_title = "Prevalence (%)",
  subtitle = "24–35 months",
  palette = 8,
  direction = -1
)

figb

```


Concentration curve for full vaccination by cumulative percentage of the population ranked by wealth index

Data preparation step

```{r}
#| label: concentration-curves-data-prep-23_35
#| message: false
#| warning: false
#| cache: false
#| include: true

cc_df_immunn_24_35 <- df_immunn_24_35 |>   
    filter(!is.na(fully_vacc), !is.na(wealth_cont), !is.na(wt)) |>     
    arrange(wealth_cont) |> 
    mutate(
    pop_rank = cumsum(wt) - wt / 2, # midpoint rank within own weight mass
    pop_rank = pop_rank / sum(wt)) |> 
  group_by(wealth_cont) |>   # collapse exact ties
  mutate(pop_rank = mean(pop_rank)) |> # give tied group its average rank
  ungroup() |> 
  arrange(pop_rank) |> 
  mutate(
        cum_prop = cumsum(wt) / sum(wt),
        cum_full_vacc = cumsum(fully_vacc*wt) / sum(fully_vacc*wt))


cc_plot_data_immunn_24_35 <- bind_rows(
  data.frame(cum_prop = 0, cum_full_vacc = 0),
  cc_df_immunn_24_35 |> select(cum_prop, cum_full_vacc)) # Adding (0, 0) for plotting

```

Generating concentration index and plotting

```{r}
#| label: fig-concentration-curves-plotting-immunn-24_35
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Concentration curves showing "

## Concentration index
ci_result_immunn_24_35 <- ci(
  ineqvar = df_immunn_24_35$wealth_cont,
  outcome = df_immunn_24_35$fully_vacc,
  weights = df_immunn_24_35$wt,
  type    = "CIw")

cc_plot_data_immunn_24_35 |> 
  ggplot(aes(x = cum_prop, y = cum_full_vacc)) +
  geom_line(color = "lightblue", linewidth = 1) +
  geom_abline(
    slope = 1, intercept = 0,
     color = "#813134") +
  annotate(
    "text",
    x = 0.05, y = 0.95,
    label = paste0(
      "Concentration index (95% CI) = ",
      round(ci_result_immunn_24_35$concentration_index, 2),
      " (",
      round(confint(ci_result_immunn_24_35)[[1]], 2), ",",
      round(confint(ci_result_immunn_24_35)[[2]], 2), ")"),
    family = "Candara",
    hjust = 0, vjust = 1,
    size = 4.0) +
  labs(
    x = "Cumulative % of children, ranked by wealth (poorest \u2192 richest)",
    y = "Cumulative % of non fully vaccinated children",
    subtitle = "24-35 months") +
  theme_classic(base_family = "Candara", base_size = 13) +
  theme(
    panel.grid = element_blank())
```




## Skilled birth attendance

Skilled birth attendance (SBA) analysis was restricted to children aged 0–35 months. In the 2022 KDHS, delivery-assistance (SBA) questions were asked only about a mother's most recent birth, and only if that child was aged 0–35 months at the time of the interview. 

SBA was defined as a delivery assisted by a doctor, nurse, or midwife/clinical officer, in line with the standard DHS/WHO definition of a trained health professional able to manage normal deliveries and identify, manage, and refer obstetric complications (KDHS 2022 variables m3a, m3b). Traditional birth attendants, relatives, and other unskilled providers were not classified as skilled.

### Distribution of skilled birth attendance at the national and subnational levels, children 0-35 months

```{r}
#| label: sba-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

df_sba <- df |>
  filter(!is.na(m3a), !is.na(m3b)) |> 
  mutate(
    sba = if_else(m3a == 1 | m3b == 1, "sba", "non sba"),
    sba_num = if_else(m3a == 1 | m3b == 1, 1, 0)) 

```

Survey setting step

```{r}
#| label: sba-svy-set
#| message: false
#| warning: false
#| cache: false
#| include: true

df_sba_survey <- svydesign(
  id = ~v021, # primary sampling unit       
  strata = ~v022, # strata     
  weights = ~wt,       
  data = df_sba,
  nest = TRUE)  

options(survey.lonely.psu = "adjust") # adjust for case where there may be 1 psu in a strata
```

Weighted prevalence of sba - national level

```{r}
#| label: weighted-prev-sba
#| message: false
#| warning: false
#| cache: false
#| include: true

tbl_sba <- tbl_svysummary(
  df_sba_survey,
  include = sba,
  statistic = all_categorical() ~ "{n} ({p})",
  digits = all_categorical() ~ c(1, 1),
  missing = "ifany",
  missing_text = "Missing",
  label = sba ~ "Skilled birth attendance")

```

Weighted prevalence of full vaccination - subnational levels

```{r}
#| label: fig-prev-counties-sba
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of skilled birth weight among caregivers of children 0-35 months at the subnationa levels. The blue line represents the national weighted prevalence"

county_sba <- svyby(
  ~ I(sba == "sba"),
  by = ~v024,
  design = df_sba_survey,
  FUN = svyciprop,
  vartype = "ci",
  na.rm = TRUE)

sba_county <- county_sba |> 
    mutate(
        county = v024,
        prevalence = `I(sba == "sba")`,
        lb = ci_l,
        ub = ci_u) |> 
    select(county, prevalence, lb, ub)


sba_county <- sba_county |> 
    mutate(
        name = names(attr(county, "labels")) |> 
            str_to_title() |> 
            str_replace_all("[[:punct:]]", " "))


# Merge the prevalence data to the shp file data
merged_sba <- kenya_shp |> 
    left_join(
        sba_county, by = c("county" = "name")) |> 
    mutate(
        across(
            c(prevalence, lb, ub), ~round(., 3)*100))

# Plotting
plot_pointrange(
    merged_sba, x_var = "county", estimate = "prevalence",
    lower = lb, upper = ub,
    reference_line = 
        as.numeric(
            str_extract(tbl_sba$table_body[3, 6][[1]],
             "(?<=\\()[0-9.]+(?=\\))")),
    y_limits = c(30, 100),
    y_breaks = seq(30, 100, 10),
    y_label = "Weighted prevalence (%)")

```

Spatial distribution of skilled birth attendance at the subnational level

```{r}
#| label: fig-prev-maps-sba
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Spatial distribution of skilled birth attendance at the subnational level"

plot_map(
  merged_sba,
  fill = prevalence,
  breaks = c(50, 70, 90, 100),
  labels = c("50-69.9", "70-89.9", "90-100"),
  legend_title = "Prevalence (%)",
  subtitle = NULL,
  palette = 8,
  direction = -1
)

```


Concentration curve for sba by cumulative percentage of the population ranked by wealth index

Data preparation step

```{r}
#| label: concentration-curves-data-prep-sba
#| message: false
#| warning: false
#| cache: false
#| include: true

cc_df_sba <- df_sba |>   
    filter(!is.na(sba_num), !is.na(wealth_cont), !is.na(wt)) |>     
    arrange(wealth_cont) |> 
    mutate(
    pop_rank = cumsum(wt) - wt / 2, # midpoint rank within own weight mass
    pop_rank = pop_rank / sum(wt)) |> 
  group_by(wealth_cont) |>   # collapse exact ties
  mutate(pop_rank = mean(pop_rank)) |> # give tied group its average rank
  ungroup() |> 
  arrange(pop_rank) |> 
  mutate(
        cum_prop = cumsum(wt) / sum(wt),
        cum_sba = cumsum(sba_num*wt) / sum(sba_num*wt))


cc_plot_data_sba <- bind_rows(
  data.frame(cum_prop = 0, cum_full_vacc = 0),
  cc_df_sba |> select(cum_prop, cum_sba)) # Adding (0, 0) for plotting

```

Generating concentration index and plotting

```{r}
#| label: fig-concentration-curves-plotting-sba
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Concentration curves showing "

## Concentration index
ci_result_sba <- ci(
  ineqvar = df_sba$wealth_cont,
  outcome = df_sba$sba_num,
  weights = df_sba$wt,
  type    = "CIw")

cc_plot_data_sba |> 
  ggplot(aes(x = cum_prop, y = cum_sba)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_abline(
    slope = 1, intercept = 0,
     color = "#813134") +
  annotate(
    "text",
    x = 0.05, y = 0.95,
    label = paste0(
      "Concentration index (95% CI) = ",
      round(ci_result_sba$concentration_index, 2),
      " (",
      round(confint(ci_result_sba)[[1]], 2), ",",
      round(confint(ci_result_sba)[[2]], 2), ")"),
    family = "Candara",
    hjust = 0, vjust = 1,
    size = 4.0) +
  labs(
    x = "Cumulative % of children, ranked by wealth (poorest \u2192 richest)",
    y = "Cumulative % of skilled birth attendance") +
  theme_classic(base_family = "Candara", base_size = 13) +
  theme(
    panel.grid = element_blank())

```
















