---
title: "How Kenyan maternal/child indicators (stunting, skilled birth attendance, immunisation) vary by region and wealth quintile"
author: Kevinson Mwangi, Elphas Abok, Jared Onsomu, and John Andrew
date: today
date-format: "DD MMMM YYYY"
format: 
  typst:
    citeproc: true
    template-partials: 
      - typst-show.typ
      - typst-template.typ
editor_options: 
  chunk_output_type: console
---

### Model Based Geostatistics - Full vaccination
#### 1. Spatial distribution of full vaccination

Packages and data 
```{r}
#| label: package-setup
#| message: false
#| warning: false
#| cache: false
#| include: true


pacman::p_load(
  tidyverse, haven, labelled, epiDisplay, survey,
  gtsummary, extrafont, flextable, PrevMap)

## Loading the data as cleaned from the descriptive and exploratory analysis

df_immunisation_mbg <- 
    read_dta(
        "./Data/Derived_data/df_immunisation.dta") |> 
          mutate(
            fully_vacc = ifelse(
              fully_vaccinated == "Fully vaccinated", 1, 0))
```


Covariates assembly

```{r}
#| label: covariates
#| message: false
#| warning: false
#| cache: false
#| include: true

### Child age
skimr::skim(df_immunisation_mbg$age_months)

### Age categorised
df_immunisation_mbg <- df_immunisation_mbg |> 
mutate(
    age_months_cat = 
      case_when(
        age_months >= 12 & age_months < 24 ~ "12-23 months",
        age_months >= 24 & age_months < 36 ~ "24-35 months",
        TRUE ~ NA_character_) |> 
      factor(
        levels = 
          c("12-23 months", "24-35 months")) |> 
      relevel(ref = "12-23 months"))

tab1(df_immunisation_mbg$age_months_cat, graph = F)

### Child gender
df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    child_sex = b4,
    child_sex = 
      factor(
        child_sex,
        levels = c(1, 2),
        labels = c("Male", "Female")))

tab1(df_immunisation_mbg$child_sex, graph = F)

### Parity
skimr::skim(df_immunisation_mbg$v201)
df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    parity = ifelse(v201 < 3, "1 or 2", "3+") |> 
      factor(
        levels = c("1 or 2", "3+")) |> 
      relevel(ref = "1 or 2"))

tab1(df_immunisation_mbg$parity, graph = F)

### Mother's age
skimr::skim(df_immunisation_mbg$v012)

df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    mother_age = v012,
    mother_age_cat = 
      case_when(
        mother_age < 20 ~ "15-19",
        mother_age >= 20 & mother_age < 35 ~ "20-34",
        mother_age >= 35 & mother_age < 50 ~ "35+",
        TRUE ~ NA_character_) |> 
      factor(
        levels = c("15-19", "20-34", "35+")) |> 
      relevel(ref = "15-19"))

tab1(df_immunisation_mbg$mother_age_cat, graph = F)

## Mother's marital status
df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    marital_status = 
      case_when(
        v501 %in% c(0, 3, 4, 5) ~ "Not in a union",
        v501 %in% c(1, 2) ~ "In a union",
        TRUE ~ NA_character_) |> 
      as.factor() |> 
      relevel(ref = "Not in a union"))

tab1(df_immunisation_mbg$marital_status, graph = F)

## Mother's education
df_immunisation_mbg |> dplyr::select(sreduc, v106, v149)

df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    education = 
      case_when(
        v106 == 0 ~ "No education",
        v106 == 1 ~ "Primary",
        v106 %in% c(2, 3) ~ "Secondary or higher",
        TRUE ~ NA_character_) |> 
      factor(
        levels = c("No education", "Primary", "Secondary or higher")) |> 
      relevel(ref = "No education"))

tab1(df_immunisation_mbg$education, graph = F)

## Mother's employment status
df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    mother_working = ifelse(v714 == 1, "Yes", "No") |> 
      as.factor() |> 
      relevel(ref = "No"))
tab1(df_immunisation_mbg$mother_working, graph = F)

## Household wealth index
df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    wealth_status = 
      case_when(
        v190 %in% c(1, 2) ~ "Low",
        v190 == 3 ~ "Middle",
        v190 %in% c(4, 5) ~ "High") |> 
      factor(
        levels = c("Low", "Middle", "High")) |> 
      relevel(ref = "Low"))

tab1(df_immunisation_mbg$wealth_status, graph = F)

## Area of residence
df_immunisation_mbg <- df_immunisation_mbg |> 
  mutate(
    residence = ifelse(v025 == 1, "Urban", "Rural") |> 
      as.factor() |> 
      relevel(ref = "Rural"))

tab1(df_immunisation_mbg$residence, graph = F)

```


Selecting the cleaned variables


```{r}
#| label: final-data
#| message: false
#| warning: false
#| cache: false
#| include: true

df_immunisation_mbg_complete <- df_immunisation_mbg |> 
  dplyr::select(
    v001, id = v021, strata = v022, wt, fully_vaccinated,
    fully_vacc, age_months, age_months_cat, child_sex,
    parity, mother_age, mother_age_cat, marital_status,
    education,   mother_working, wealth_status, residence)

```


Survey setting


```{r}
#| label: survey-setting
#| message: false
#| warning: false
#| cache: false
#| include: true

mysurvey <- svydesign(
  id = ~id,       
  strata = ~strata,      
  weights = ~wt,       
  data = df_immunisation_mbg_complete,
  nest = TRUE)  

options(survey.lonely.psu = "adjust") # adjust for case where there may be 1 psu in a strata

```


Characteristics of study participants


```{r}
#| label: tbl-participant-characteristics
#| message: false
#| warning: false
#| cache: false
#| include: false

gtsummary::theme_gtsummary_journal("jama")

tbl_svysummary(
  mysurvey,
  by = fully_vaccinated,
  include = c(
    age_months_cat, child_sex,
    mother_age_cat, marital_status,
    education, mother_working, parity, wealth_status, residence),
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    all_categorical() ~ "{n} ({p})"),
  digits = list(
    all_continuous() ~ c(2, 1),
    all_categorical() ~ c(0, 1)),
  label = list(
    age_months_cat ~ "Child age (months)",
    child_sex ~ "Child sex",
    mother_age_cat ~ "Age of the mother (years)",
    marital_status ~ "Mother's marital status",
    education ~ "Highest level of completed education of mother",
    mother_working ~ "Mother's employment status",
    parity ~ "Parity",
    wealth_status ~ "Household wealth status",
    residence ~ "Area of residence"),
  missing = "ifany",
  missing_text = "Missing") |>
  add_overall() |>
  add_p(
    test = list(
      all_continuous() ~ "svy.t.test",
      all_categorical() ~ "svy.chisq.test"),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)) |>
  italicize_levels() |> 
  as_flex_table() |> 
  autofit() |> 
  flextable::font(fontname = "Candara", part = "all") |> 
  flextable::fontsize(size = 11, part = "all") |> 
  flextable::align(
    j = c(2, 3, 4, 5), 
    align = c("left", "left", "left", "right"), 
    part = "all")
  
```


Univariable regression models


```{r}
#| label: univariable-regression
#| message: false
#| warning: false
#| cache: false
#| include: false

tbl_uvregression(
  mysurvey,
  method = svyglm,
  y = fully_vacc,
  method.args = list(family = quasibinomial()),  
  include = c(age_months_cat, child_sex, parity,
              mother_age_cat, marital_status,
              education, mother_working, wealth_status, residence),
  exponentiate = TRUE,      
  label = list(
    age_months_cat ~ "Child age (months)",
    child_sex ~ "Child sex",
    mother_age_cat ~ "Age of the mother (years)",
    marital_status ~ "Mother's marital status",
    education ~ "Highest level of completed education of mother",
    mother_working ~ "Mother's employment status",
    parity ~ "Parity",
    wealth_status ~ "Household wealth status",
    residence ~ "Area of residence"),
  pvalue_fun = ~ style_pvalue(.x, digits = 3)) |> 
  italicize_levels() |> 
  bold_p() |> 
   as_flex_table() |> 
  autofit() |> 
  flextable::font(fontname = "Candara", part = "all") |> 
  flextable::fontsize(size = 11, part = "all") |> 
  align(
    j = c(2, 3, 4), 
    align = c("left", "left", "right"), 
    part = "all")
```



```{r}
#| label: multivariable-regression
#| message: false
#| warning: false
#| cache: false
#| include: false

mult_model_immunisation <- tbl_regression(
  svyglm(
  fully_vacc ~ age_months_cat + child_sex +
    mother_age_cat + marital_status + education + 
    mother_working + parity + wealth_status + residence,
  design = mysurvey,
  family = quasibinomial()),
  exponentiate = TRUE,
  label = list(
    age_months_cat ~ "Child age (months)",
    child_sex ~ "Child sex",
    mother_age_cat ~ "Age of the mother (years)",
    marital_status ~ "Mother's marital status",
    education ~ "Highest level of completed education of mother",
    mother_working ~ "Mother's employment status",
    parity ~ "Parity",
    wealth_status ~ "Household wealth status",
    residence ~ "Area of residence"),
  pvalue_fun = ~ style_pvalue(.x, digits = 3)) |>
  italicize_levels() |>
  bold_p() |>
  as_flex_table() |>
  autofit() |>
  font(fontname = "Candara", part = "all") |>
  fontsize(size = 11, part = "all") |>
  align(
    j = c(2, 3),
    align = c("left", "right"),
    part = "all")

mult_model_immunisation

write_csv(
mult_model_immunisation$body$dataset |> 
  as.data.frame() |> 
  filter(
    label %in% 
      c("Household wealth status", "Low", "Middle", "High")),

    "./Data/Derived_estimates/dhs_immunisation_WI_adjusted_OR.csv")

```



```{r}
#| label: recode-selected-variables
#| message: false
#| warning: false
#| cache: false
#| include: false

df_immunisation_mbg_complete_final <- df_immunisation_mbg_complete |> 
  mutate(
    age_months_cat_rec = 
      ifelse(
    age_months_cat == "12-23 months", 0, 1),
    child_sex_rec = ifelse(child_sex == "Male", 0, 1),
    parity_rec = ifelse(parity == "1 or 2", 1, 0),
    age_mother_rec = ifelse(mother_age_cat == "35+", 1, 0),
    marital_status_rec = ifelse(marital_status == "In a union", 1, 0),
    education_rec = ifelse(education == "No education", 0, 1),
    employment_rec = ifelse(mother_working == "Yes", 1, 0),
    wealth_status_rec = ifelse(wealth_status == "Low", 0, 1),
    residence_rec = ifelse(residence == "Rural", 0, 1))

```


Data collapse in prep for MBG


```{r}
#| label: data-collapse
#| message: false
#| warning: false
#| cache: false
#| include: false

collapsed_immunisation_data <- df_immunisation_mbg_complete_final |>
  group_by(v001) |>
  summarise(
    fully_vaccinated = sum(fully_vacc, na.rm = TRUE),
    total = sum(!is.na(fully_vacc)),
    across(
      c(
        age_months_cat_rec, child_sex_rec, parity_rec,
        age_mother_rec, marital_status_rec, education_rec,
        employment_rec, wealth_status_rec, residence_rec),
      ~ mean(.x, na.rm = TRUE)),
    .groups = "drop")

```


MBG Exploratory analysis

```{r}
#| label: exploratory-analysis
#| message: false
#| warning: false
#| cache: false
#| include: false

## Log logit

collapsed_immunisation_data$logit <- log(
  (collapsed_immunisation_data$fully_vaccinated + 0.5) / (collapsed_immunisation_data$total - collapsed_immunisation_data$fully_vaccinated + 0.5))


theme_set(
  theme_bw(base_family = "Candara", base_size = 14) +
    theme(
      panel.grid = element_blank()))

## Child age
age_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = age_months_cat_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Children with older age", 
    y = "Empirical logit")


## Child sex
sex_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = child_sex_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Gender", 
    y = "Empirical logit")


## Parity
parity_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = parity_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Parity", 
    y = "Empirical logit")


## Mother age
age_mother_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = age_months_cat_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Older mothers", 
    y = "Empirical logit")


## Marital status
marital_status_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = marital_status_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Married women", 
    y = "Empirical logit")

## Education
education_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = education_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Higher education coverage", 
    y = "Empirical logit")


## Employment
employment_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = employment_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Coverage of employment", 
    y = "Empirical logit")


## Wealth status
wealth_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = wealth_status_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Coverage of wealth status", 
    y = "Empirical logit")


## Residence
residence_graph <- collapsed_immunisation_data |> 
  ggplot(
    aes(x = residence_rec, y = logit)) + 
  geom_point() + 
  stat_smooth(
    method = "gam",
    formula = y ~ s(x), 
    se = F) +
  stat_smooth(
    method = "lm",
    formula = y ~ x,
    col = "red",
    lty = 2, 
    se = F) + 
  labs(
    x = "Coverage of residence", 
    y = "Empirical logit")


library(patchwork)

(age_graph / sex_graph / parity_graph / age_mother_graph) | (marital_status_graph / education_graph / employment_graph) | (wealth_graph / residence_graph)

```


Checking for multicollinearity


```{r}
#| label: correlation-plot
#| message: false
#| warning: false
#| cache: false
#| include: false

library(ggcorrplot)

gg_cor_immunisation <- ggcorrplot(
  cor(
    collapsed_immunisation_data[, c(
      "age_months_cat_rec", "child_sex_rec", "parity_rec",
        "age_mother_rec", "marital_status_rec", "education_rec",
        "employment_rec", "wealth_status_rec", "residence_rec")],
    use = "pairwise.complete.obs"),
  type = "lower",
  show.legend = F,
    ggtheme = ggplot2::theme_bw(base_family = "Candara") +
         theme(panel.grid = element_blank()),
    hc.order = F,
    show.diag = F,
    outline.col = "white",
    lab = TRUE,
    legend.title = "Correlation")

gg_cor_immunisation

```


Grid formation


```{r}
#| label: spatial-data-prep
#| message: false
#| warning: false
#| cache: false
#| include: false

pacman::p_load(rgeoboundaries, sf)

kenya_shp <- geoboundaries("Kenya", "adm1") |> 
  st_as_sf()

write_sf(kenya_shp, "./Data/Derived_data/kenya_counties.shp")


## DHS cluster shapefile

cluster_df <- st_read("./Data/Shapefiles/KEGE8AFL.shp")
  
# Merging the collapsed data to the cluster shp
collapsed_immunisation_data$v001 <- 
  as.integer(collapsed_immunisation_data$v001)

cluster_df$DHSCLUST <- as.integer(cluster_df$DHSCLUST)

immun_data_shp <- cluster_df |> 
  left_join(collapsed_immunisation_data, by = c("DHSCLUST" = "v001"))

immun_data_shp_utm <- st_transform(immun_data_shp, 21037)


## Extract UTM coordinates
utm_coords <- st_coordinates(immun_data_shp_utm)
immun_data_shp_utm$utm_x <- utm_coords[, 1]
immun_data_shp_utm$utm_y <- utm_coords[, 2]


## Extract longitude/latitude
immun_data_shp_latlon <- st_transform(immun_data_shp_utm, 4326)
latlon_coords <- st_coordinates(immun_data_shp_latlon)

immun_data_shp_utm$longitude <- latlon_coords[, 1]
immun_data_shp_utm$latitude <- latlon_coords[, 2]

save(
  immun_data_shp_utm,
  file = "./Data/Derived_data/immun_data_shp_utm.RData")


## Kenya Grid creation
kenya_shp2 <- st_as_sf(kenya_shp)

## Transform shapefile
kenya_shp_utm <- st_transform(kenya_shp2, 21037)
utm_coords <- st_coordinates(immun_data_shp_utm)
immun_data_shp$utm_x <- utm_coords[, 1]
immun_data_shp$utm_y <- utm_coords[, 2]

## Union geometry
kenya_union <- st_union(kenya_shp_utm)

## Create grid centers
kenya.grid.sq <- st_make_grid(
  kenya_shp_utm,
  cellsize = 5000,
  what = "centers")


## Keep only points inside Kenya
kenya.inout <- st_intersects(
  kenya.grid.sq,
  kenya_union,
  sparse = F)

kenya.grid <- kenya.grid.sq[kenya.inout]


## Convert grid to sf object
kenya.grid.sf <- st_sf(
  geometry = kenya.grid,
  crs = 21037)

## Join county names from polygons
kenya.grid.sf <- st_join(
  kenya.grid.sf,
  kenya_shp_utm["shapeName"],
  join = st_intersects)


## Rename county column
kenya.grid.sf <- kenya.grid.sf %>%
  rename(county = shapeName)


# Extract UTM coordinates
utm_coords <- st_coordinates(kenya.grid.sf)

# Transform to lon/lat
kenya.grid.latlon <- st_transform(kenya.grid.sf, 4326)

# Extract lon/lat coordinates
latlon_coords <- st_coordinates(kenya.grid.latlon)

# Final dataframe
kenya.grid.pred <- kenya.grid.sf %>%
  st_drop_geometry() %>%
  mutate(
    utm_x    = utm_coords[,1],
    utm_y    = utm_coords[,2],
    longitude = latlon_coords[,1],
    latitude  = latlon_coords[,2])

```


Empirical logit


```{r}
#| label: empirical-logit
#| message: false
#| warning: false
#| cache: false
#| include: true

## Empirical logit variable
immun_data_shp_utm$fully_vaccination_prev2 <- 
  immun_data_shp_utm$fully_vaccinated / immun_data_shp_utm$total *100

plot(immun_data_shp_utm["fully_vaccination_prev2"])

library(tmap)

tm_shape(immun_data_shp_utm) +
  tm_symbols(
    size = "fully_vaccination_prev2",
    col = "fully_vaccination_prev2",
    palette = "viridis",
    alpha = 0.7,
    border.col = "black",
    border.lwd = 0.2) +
  tm_layout(legend.outside = TRUE)

immunisation_df <- data.frame(immun_data_shp_utm)

immunisation_df$logit <- log(
  (immunisation_df$fully_vaccinated + 0.5) / 
    (immunisation_df$total - immunisation_df$fully_vaccinated + 0.5))

```


\newpage

Residual spatial correlation

```{r}
#| label: residual-spatial-correlation
#| message: false
#| warning: false
#| cache: false
#| include: false

## Variogram

library(PrevMap)

immunisation_df$utm_x <- immunisation_df$utm_x/1000
immunisation_df$utm_y <- immunisation_df$utm_y/1000

vars <- c(
  "fully_vaccinated", "age_months_cat_rec", "child_sex_rec",
  "parity_rec", "age_mother_rec", "marital_status_rec",
  "education_rec", "employment_rec", "wealth_status_rec",
  "residence_rec", "total", "utm_x", "utm_y", "longitude",  
  "latitude")


immunisation_df_clean <- immunisation_df %>%
  dplyr::select(all_of(vars)) %>%
  na.omit()

par(bty = "o")


variog_immunisation <- spat.corr.diagnostic(
  fully_vaccinated ~ age_months_cat_rec + child_sex_rec + 
    parity_rec + age_mother_rec + marital_status_rec +
    education_rec + employment_rec + wealth_status_rec +
    residence_rec,
  units.m = ~ total,
  coords = ~utm_x + utm_y,
  likelihood = "Binomial",
  lse.variogram = T,
  data = immunisation_df_clean)



pars   <- variog_immunisation$lse.variogram
sigma2 <- pars["sigma^2"]
phi    <- pars["phi"]
nugget <- pars["tau^2"]

variog_df_immunisation <- data.frame(
  distance = variog_immunisation$distance.bins,
  obs      = variog_immunisation$obs.variogram,
  lower    = variog_immunisation$lower.lim,
  upper    = variog_immunisation$upper.lim) |>
  mutate(
    fitted = nugget + sigma2 * (1 - exp(-distance / phi)))

p_variog_immunisation <- 
  ggplot(variog_df_immunisation, aes(x = distance)) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper),
     fill = "grey80", alpha = 0.5) +
  geom_line(
    aes(y = obs), color = "black", linewidth = 0.8) +
  geom_point(
    aes(y = obs), color = "black", size = 2) +
  geom_line(
    aes(y = fitted), color = "blue", 
    linetype = "dashed", linewidth = 0.8) +
  labs(
    x = "Spatial distance",
    y = "Variogram",
    subtitle = "Full vaccination") +
  theme_bw(base_family = "Candara", base_size = 14) + 
  theme(
    plot.subtitle = element_text(hjust = .5),
    panel.grid = element_blank())

p_variog_immunisation

```

Covariates standardisation

```{r}
#| label: covariate-standardisation
#| message: false
#| warning: false
#| cache: false
#| include: true

for (i in c("age_months_cat_rec", "child_sex_rec", "parity_rec",
"age_mother_rec", "marital_status_rec", "education_rec",      "employment_rec", "wealth_status_rec", "residence_rec")) {

 newvar <- paste0(i, "_z")
 immunisation_df_clean[[newvar]] <- 
  as.numeric(scale(immunisation_df_clean[[i]]))
}

```

Unique spatial locations

```{r}
#| label: unique-spatial-locations
#| message: false
#| warning: false
#| cache: false
#| include: true

ID.coords <- create.ID.coords(
  data = immunisation_df_clean, 
  ~I(utm_x)+I(utm_y))
set.seed(123)

```

