---
title: "Task 2: How Kenyan maternal/child indicators (stunting, skilled birth attendance, immunisation) vary by region and wealth quintile."
author: "Kevinson Mwangi & Elphas Abok"
date: today
format: html
---

## Loading the KDHS Data 2022 (KR Record)
```{r}
#| label: setup
#| message: false
#| warning: false
#| cache: false
#| include: true

rm(list = ls())
pacman::p_load(
  tidyverse, haven, labelled, epiDisplay, survey,
  gtsummary, extrafont, flextable)

df <- read_dta("../Mini project/Data/KEKR8BDT/KEKR8BFL.DTA")

```

## Data preparation
```{r}
#| label: data-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

## Restrict the analysis to the youngest alive child, living with the responded 
df <- df |>  
    mutate(
        age_months = v008 - b3) |> 
    filter(
        b5 == 1, # child is alive
        b9 == 0, # child lives with the respondent
        !is.na(v003) # mother's line number must be available
        ) |>
    group_by(
        v001, 
        v002, 
        v003 # grouping by cluster, household, and mothers line number
    ) |> 
    mutate(
        bidxr = 
            rank(bidx, ties.method = "first")) |> 
    filter(
        bidxr == 1 # Keeping the youngest child 
        ) |> 
    ungroup()

```

The datasets includes a total of `r dim(df)[[1]]` children from `r length(unique(df$v001))` clusters across Kenya. 

## The ouctomes of interest
### 1. Stunting

```{r}
#| label: fig-stunting
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Histogram of HAZ scores"

df_stunting <- df |> 
    filter(
        age_months > 5 & age_months < 60) |> 
  mutate(
    hw70 = ifelse( #hw70 is the height-for-age zscore and should be divided by 100
        hw70 >
         1000, NA, hw70),
    haz = hw70/100,
    
    stunted = ifelse(haz < -2, "Stunted", "Not stunted"))
 
df_stunting |> 
    ggplot(
        aes(x = haz)) +
            geom_histogram(color = "white") +
            labs(
                x = "Height-for-age z-score (HAZ scores)",
                y = "Frequency") + 
    theme_bw(base_family = "Candara", base_size = 14) +
    theme(panel.grid = element_blank())
```

Stunting was defined as height-for-age z-score (HAZ) < -2 based on the WHO 2006 growth reference standards. The analysis included a total of `r dim(df_stunting)[[1]]`. According to the WHO guidelines, HAZ scores range from -6 to 6 and this was ascertained for the data as shown in @fig-stunting.

\newpage

## Immunisation, restricted to children 12-23 months
Immunisation analysis was restricted to only 12-23 months. Full vaccination was defined as having taken BCG, HepB (birth dose), 3 doses of DPT-HepB-Hib, 3 [4] doses of OPV, 1 dose of IPV, 3 [2] doses of pneumococcal vaccine, 3 [2] doses of rotavirus vaccine, and 1 dose of MCV, according to the national vaccination schedule.




