---
title: "Task 2: How Kenyan maternal/child indicators (stunting, skilled birth attendance, immunisation) vary by region and wealth quintile."
author: "Kevinson Mwangi & Elphas Abok"
date: today
format: html
---

## Loading the KDHS Data 2022 (KR Record)
```{r}
#| label: setup
#| message: false
#| warning: false
#| cache: false
#| include: true

rm(list = ls())
pacman::p_load(
  tidyverse, haven, labelled, epiDisplay, survey,
  gtsummary, extrafont, flextable, rKenyaCensus, sf)

df <- read_dta("../Mini project/Data/KEKR8BDT/KEKR8BFL.DTA")

```

## Data preparation
```{r}
#| label: data-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

## Restrict the analysis to the youngest alive child, living with the responded 
df <- df |>  
    mutate(
        age_months = v008 - b3) |> 
    filter(
        b5 == 1, # child is alive
        b9 == 0, # child lives with the respondent
        !is.na(v003) # mother's line number must be available
        ) |>
    group_by(
        v001, 
        v002, 
        v003 # grouping by cluster, household, and mothers line number
    ) |> 
    mutate(
        bidxr = 
            rank(bidx, ties.method = "first")) |> 
    filter(
        bidxr == 1 # Keeping the youngest child 
        ) |> 
    ungroup()

```

The datasets includes a total of `r dim(df)[[1]]` children from `r length(unique(df$v001))` clusters across Kenya. 

## The ouctomes of interest
### Stunting
#### National weighted n and prevalence of stunting

```{r}
#| label: fig-stunting
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Histogram of HAZ scores"

df_stunting <- df |> 
    filter(
        age_months > 5 & age_months < 60) |> 
  mutate(
    hw70 = ifelse( #hw70 is the height-for-age zscore and should be divided by 100
        hw70 >
         1000, NA, hw70),
    haz = hw70/100,
    
    stunted = ifelse(haz < -2, "Stunted", "Not stunted"))
 
df_stunting |> 
    ggplot(
        aes(x = haz)) +
            geom_histogram(color = "white") +
            labs(
                x = "Height-for-age z-score (HAZ scores)",
                y = "Frequency") + 
    theme_bw(base_family = "Candara", base_size = 14) +
    theme(panel.grid = element_blank())
```


```{r}
#| label: survey-setting
#| message: false
#| warning: false
#| cache: false
#| include: true

df_stunting <- df_stunting |> 
    mutate(
        wt = v005/1000000 # sample weight
    )
# Survey setting
stunting_survey <- svydesign(
  id = ~v021, # primary sampling unit       
  strata = ~v022, # strata     
  weights = ~wt,       
  data = df_stunting,
  nest = TRUE)  

options(survey.lonely.psu = "adjust") # adjust for case where there may be 1 psu in a strata

# Weighted prevalence of stunting
tbl <- tbl_svysummary(
  stunting_survey,
  include = stunted,
  statistic = all_categorical() ~ "{n} ({p})",
  digits = all_categorical() ~ c(1, 1),
  missing = "ifany",
  missing_text = "Missing",
  label = stunted ~ "Stunting status")
```

Stunting was defined as height-for-age z-score (HAZ) < -2 based on the WHO 2006 growth reference standards. The analysis included a total of `r dim(df_stunting)[[1]]`. According to the WHO guidelines, HAZ scores range from -6 to 6 and this was ascertained for the data as shown in @fig-stunting. The data was then survey-set to account for the two-stage sampling design of DHS. The weighted n (prevalence %) of stunting among the 6-59 months old children was `r tbl$table_body[3, 6][[1]]`.


#### Weighted prevalence of stunting at the subnational levels

```{r}
#| label: fig-prev-counties
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of stunting among children 6-59 months at the subnationa levels. The blue line represents the national weighted prevalence"

county_stunting <- svyby(
  ~ I(stunted == "Stunted"),
  by = ~v024,
  design = stunting_survey,
  FUN = svyciprop,
  vartype = "ci",
  na.rm = TRUE)

stunting_county <- county_stunting |> 
    mutate(
        county = v024,
        prevalence = `I(stunted == "Stunted")`,
        lb = ci_l,
        ub = ci_u) |> 
    select(county, prevalence, lb, ub)


stunting_county <- stunting_county |> 
    mutate(
        name = names(attr(county, "labels")) |> 
            str_to_title() |> 
            str_replace_all("[[:punct:]]", " "))


# Fetching Kenya counties shapefiles
kenya_shp <- rKenyaCensus::KenyaCounties_SHP |> 
    janitor::clean_names() |> 
    st_as_sf() |> 
    select(county) |> 
    mutate(
        county = str_to_title(county) |> 
            str_replace_all("[[:punct:]]", " "))

kenya_shp[kenya_shp$county == "Nairobi City", "county"] <- "Nairobi"

# Merge the prevalence data to the shp file data
merged_stunting <- kenya_shp |> 
    left_join(
        stunting_county, by = c("county" = "name")) |> 
    mutate(
        across(
            c(prevalence, lb, ub), ~round(., 3)*100))

# Plotting the prevalence across the counties
merged_stunting |> 
    ggplot(
        aes(
            x = fct_reorder(county, prevalence, .desc = TRUE),
            y = prevalence)) +
    geom_pointrange(
        aes(ymin = lb, ymax = ub),
        size = .3,
        color = "gray30") +
    scale_y_continuous(
        limits = c(5, 50),
        breaks = seq(5, 50, 5)) +
    geom_hline(
        yintercept = as.numeric(
            str_extract(tbl$table_body[3, 6][[1]], "(?<=\\()[0-9.]+(?=\\))")),
        lty = 2,
        color = "blue") +
    labs(
        x = NULL,
        y = "Weighed prevalence (%)") + 
    theme_bw(base_family = "Candara", base_size = 14) +
    theme(
        axis.text = element_text(color = "black"),
        axis.text.x = element_text(angle = 90, vjust = .5)) 
```


@fig-prev-counties shows the weighted prevalence of stunting among children 6-59 months at the subnational levels with the blue dotted line showing the weighted national prevalence.

```{r}
#| label: fig-prev-map
#| message: false
#| warning: false
#| cache: false
#| include: true
#| fig-cap: "Weighted prevalence of stunting among children 6-59 months at the subnationa levels"
        
## Prevalence maps
merged_stunting |> 
    mutate(
        prevalence_bin = cut(
            prevalence,
            breaks = c(0, 10, 20, 30, 40, 50),
            labels = c("0–10", "11–20", "21–30", "31–40", "41–50"),
            include.lowest = TRUE)) |> 
    ggplot() + 
    geom_sf(aes(fill = prevalence_bin)) + 
    scale_fill_brewer(palette = 6) +
    labs(
        fill = "Prevalence (%)") + 
    theme_minimal(base_family = "Candara", base_size = 14) +
    theme(
        axis.text = element_blank(),
        legend.background = element_rect(color = "gray70"),
        panel.grid = element_blank()) 
    
```


@fig-prev-map shows the weighted prevalence of stunting among children 6-59 months at the subnational levels. Darker shades shows increasing prevalence


## Immunisation, restricted to children 12-23 months
Immunisation analysis was restricted to only 12-23 months. Full vaccination was defined as having taken BCG, HepB (birth dose), 3 doses of DPT-HepB-Hib, 3 [4] doses of OPV, 1 dose of IPV, 3 [2] doses of pneumococcal vaccine, 3 [2] doses of rotavirus vaccine, and 1 dose of MCV, according to the national vaccination schedule.

#### National weighted n and prevalence of stunting
```{r}

#| label: immunisation-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

df_immunn_12_23 <- df |> 
    filter(
        age_months >= 12 & age_months < 24) |> 
    mutate(
        fully_vaccinated = 
            ifelse(
            h2 %in% c(1, 2, 3) & # one dose of BCG vaccine

            h0 %in% c(1, 2, 3) & # Oral polio vaccine (birth dose)
            h4 %in% c(1, 2, 3) & # Oral polio vaccine 1
            h6 %in% c(1, 2, 3) & # Oral polio vaccine 2
            h8 %in% c(1, 2, 3) & # Oral polio vaccine 3
            h60 %in% c(1, 2, 3) & # one dose of inactivated polio vaccine

            

            )

    )

```


