---
title: "How Kenyan maternal/child indicators (stunting, skilled birth attendance, immunisation) vary by region"
author: Kevinson Mwangi, Elphas Abok, Jared Onsomu, and John Andrew
date: today
date-format: "DD MMMM YYYY"
format: 
  typst:
    citeproc: true
    template-partials: 
      - typst-show.typ
      - typst-template.typ
editor_options: 
  chunk_output_type: console
---

# SMOOTHED GEOSTATISTICAL MAP OF CHILDHOOD FULL VACCINATION PREVALENCE
# Bayesian spatial (SPDE / INLA) model, DHS cluster data, Kenya

```{r}
#| label: setup
#| message: false
#| warning: false
#| cache: false
#| include: true

pacman::p_load(tidyverse, sf, sp, INLA, terra, 
rKenyaCensus, viridis, scales, patchwork)  
```

```{r}
#| label: mesh-settings
#| message: false
#| warning: false
#| cache: false
#| include: true

grid_resolution_m <- 5000     # prediction raster cell size, metres (UTM)
mesh_max_edge <- c(10000, 50000)  # inner / outer triangle edge length
mesh_cutoff <- 5000    # minimum distance between mesh points
mesh_offset <- c(10000, 50000) # inner / outer extension buffer
exceedance_threshold <- 0.20    

```


```{r}
#| label: data-prep
#| message: false
#| warning: false
#| cache: false
#| include: true

load("./Data/Derived_data/immun_data_shp_utm.RData")

dat <- immun_data_shp_utm |>
  dplyr::select(
    DHSID, ADM1NAME, LATNUM, LONGNUM, fully_vaccinated, total, age_months_cat_rec, child_sex_rec, parity_rec, age_mother_rec, marital_status_rec, education_rec, employment_rec,
    wealth_status_rec, residence_rec,
    utm_x, utm_y, geometry) |>
  mutate(
    prevalence = fully_vaccinated / total,
    cluster_id = factor(row_number()))

coords <- cbind(dat$utm_x, dat$utm_y)

```

```{r}
#| label: quick-eda
#| message: false
#| warning: false
#| cache: false
#| include: true

# Quick EDA: raw observed cluster prevalence, before any modelling

immunisation_observed_cluster <- dat |>
  ggplot() +
  geom_sf(
    data = rKenyaCensus::KenyaCounties_SHP |> 
      st_as_sf(), 
  fill = "white") +
  geom_sf(
    aes(
      colour = prevalence),
      alpha = 0.85,
      size = .55) +
  scale_colour_viridis_c(
    name = "Prevalence", 
    option = "F",
    labels = scales::percent) +
  scale_size_continuous(
    name = "Children\nsampled",
     range = c(1, 5)) +
  labs(
    subtitle = "Full vaccination") +
  theme_minimal(base_family = "Candara", base_size = 12) +
  theme(
   axis.text = element_blank(),
   axis.title = element_blank(),
   plot.subtitle = element_text(hjust = .55, vjust = -15),
   panel.grid = element_blank(),
   plot.background = element_rect(fill = "gray96"))

immunisation_observed_cluster

ggsave(
  "./Plots/Vaccination_observed_prev_map.png",
  immunisation_observed_cluster,
  width = 8,
  height = 6,
  dpi = 500,
  bg = "white")

```  

# STUDY-AREA BOUNDARY 

```{r}
#| label: area-boundary
#| message: false
#| warning: false
#| cache: false
#| include: true

kenya <- rKenyaCensus::KenyaCounties_SHP |>
  st_as_sf() |>
  summarise() |>                 
  st_transform(st_crs(dat))      

```

# MESH  (boundary-aware, extended domain)

```{r}
#| label: area-boundary
#| message: false
#| warning: false
#| cache: false
#| include: true

kenya_boundary <- inla.sp2segment(
  sf::as_Spatial(kenya))

mesh <- inla.mesh.2d(
  loc      = coords,
  boundary = kenya_boundary,
  max.edge = mesh_max_edge,
  cutoff   = mesh_cutoff,
  offset   = mesh_offset)

plot(mesh, main = "Mesh with cluster locations")
points(coords, col = "red", pch = 16, cex = 0.5)

```

#  SPDE MODEL 

```{r}
#| label: spde-model
#| message: false
#| warning: false
#| cache: false
#| include: true

spde <- inla.spde2.pcmatern(
  mesh,
  prior.range = c(50000, 0.5),
  prior.sigma = c(1, 0.01))

index <- inla.spde.make.index(name = "spatial", n.spde = spde$n.spde)

A <- inla.spde.make.A(mesh, loc = coords)

```

# ESTIMATION STACK  
```{r}
#| label: estimation-stack
#| message: false
#| warning: false
#| cache: false
#| include: true

stack <- inla.stack(
  data = list(
    y = dat$fully_vaccinated,
    Ntrials = dat$total
  ),
  A = list(A, 1),
  effects = list(
    spatial = index,
    data.frame(
      intercept  = 1,
      wealth     = factor(dat$wealth_status_rec),
      education  = factor(dat$education_rec),
      residence  = factor(dat$residence_rec),
      child_age  = factor(dat$age_months_cat_rec),
      sex        = factor(dat$child_sex_rec),
      parity     = factor(dat$parity_rec),
      cluster_id = dat$cluster_id
    )
  ),
  tag = "est")

```


# MODEL FORMULA  

```{r}
#| label: estimation-stack
#| message: false
#| warning: false
#| cache: false
#| include: true

formula <- y ~
  0 + intercept +
  wealth +
  education +
  residence +
  child_age +
  sex +
  parity +
  f(spatial, model = spde) +
  f(cluster_id, model = "iid")   

```

# FIT THE ESTIMATION MODEL

```{r}
#| label: model-fitting
#| message: false
#| warning: false
#| cache: false
#| include: true

fit <- inla(
  formula,
  family  = "binomial",
  Ntrials = dat$total,
  data    = inla.stack.data(stack),
  control.predictor = list(
    A       = inla.stack.A(stack),
    compute = TRUE
  ),
  control.compute = list(
    dic  = TRUE,
    waic = TRUE,
    cpo  = TRUE          
  )
)

summary(fit)

```


# Diagnostics 

```{r}
#| label: diagnostics
#| message: false
#| warning: false
#| cache: false
#| include: true


cat("Number of CPO failures:", sum(fit$cpo$failure > 0, na.rm = TRUE), "\n")

spde_result <- inla.spde2.result(fit, "spatial", spde)

cat("Posterior median spatial range (km):",
    round(inla.qmarginal(0.5, spde_result$marginals.range.nominal[[1]]) / 1000, 1), "\n")

cat("Posterior median spatial sigma:",
    round(inla.qmarginal(0.5, spde_result$marginals.variance.nominal[[1]])^0.5, 3), "\n")

# Fitted vs. observed prevalence at the surveyed clusters 

idx_est <- inla.stack.index(stack, "est")$data
fitted_est <- fit$summary.fitted.values[idx_est, ]

diagnostics_immunisation <- ggplot(
  data.frame(observed = dat$prevalence, fitted = fitted_est$mean),
  aes(observed, fitted)) +
  geom_point(alpha = 0.6) +
  geom_abline(
    slope = 1, 
    intercept = 0, 
    lty = 2, 
    colour = "grey40") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    subtitle = "Fitted vs. observed cluster prevalence - Full vaccination",
    x = "Observed",
    y = "Model fitted (posterior mean)") +
  theme_bw(base_family = "Candara", base_size = 12) +
  theme(
    panel.grid = element_blank())

diagnostics_immunisation

ggsave(
  "./Plots/fitted_vs_observed_vaccination.png",
  diagnostics_immunisation,
  width = 5,
  height = 5,
  dpi = 500,
  bg = "white")

```



# PREDICTION GRID  

```{r}
#| label: prediction-grid
#| message: false
#| warning: false
#| cache: false
#| include: true
#| 
kenya_vect <- terra::vect(kenya)
r_template <- terra::rast(kenya_vect, resolution = grid_resolution_m)
r_mask <- terra::rasterize(kenya_vect, r_template, field = 1)

idx_in <- which(!is.na(terra::values(r_mask)))
coords_pred  <- terra::xyFromCell(r_mask, idx_in) 
n_pred <- nrow(coords_pred)

cat("Number of prediction grid cells inside Kenya:", n_pred, "\n")

# Projector matrix from mesh nodes to the prediction locations
A_pred <- inla.spde.make.A(mesh, loc = coords_pred)

```


# PREDICTION COVARIATES 

```{r}
#| label: prediction-covariates
#| message: false
#| warning: false
#| cache: false
#| include: true
#| 
# Individual/household-level covariates (child age, sex, parity, wealth,
# education, residence) don't have a well-defined value "at a location"
# independent of a specific household -- there's no single correct answer
# to "what is the wealth level at this pixel?". Working option described below
#
#   (a) Set every covariate to a single population-weighted
#       reference profile (the empirically most common level, weighted by
#       number of children sampled), applied UNIFORMLY across the grid.
#       This yields a "reference-child" prevalence surface: how prevalence
#       varies spatially for a child with typical characteristics. All
#       spatial pattern in the resulting map comes from the estimated
#       spatial field -- nothing is borrowed from nearby clusters, so there
#       are no artificial Voronoi-cell boundaries.


get_weighted_mode <- function(level_vec, weight_vec) {
  agg <- tapply(weight_vec, level_vec, sum, na.rm = TRUE)
  names(agg)[which.max(agg)]
}

ref_levels <- list(
  wealth    = get_weighted_mode(dat$wealth_status_rec,  dat$total),
  education = get_weighted_mode(dat$education_rec,      dat$total),
  residence = get_weighted_mode(dat$residence_rec,       dat$total),
  child_age = get_weighted_mode(dat$age_months_cat_rec,  dat$total),
  sex       = get_weighted_mode(dat$child_sex_rec,       dat$total),
  parity    = get_weighted_mode(dat$parity_rec,          dat$total)
)

print(ref_levels)

pred_data <- data.frame(
  intercept  = 1,
  wealth     = factor(rep(ref_levels$wealth,    n_pred), levels = levels(factor(dat$wealth_status_rec))),
  education  = factor(rep(ref_levels$education, n_pred), levels = levels(factor(dat$education_rec))),
  residence  = factor(rep(ref_levels$residence, n_pred), levels = levels(factor(dat$residence_rec))),
  child_age  = factor(rep(ref_levels$child_age, n_pred), levels = levels(factor(dat$age_months_cat_rec))),
  sex        = factor(rep(ref_levels$sex,       n_pred), levels = levels(factor(dat$child_sex_rec))),
  parity     = factor(rep(ref_levels$parity,    n_pred), levels = levels(factor(dat$parity_rec))))

stack_pred <- inla.stack(
  data    = list(y = NA),
  A       = list(A_pred, 1),
  effects = list(spatial = index, pred_data),
  tag     = "prediction")

stack_full <- inla.stack(stack, stack_pred)
```


# REFIT WITH THE PREDICTION STACK APPENDED


```{r}
#| label: prediction-refit
#| message: false
#| warning: false
#| cache: false
#| include: true

fit_pred <- inla(
  formula,
  family  = "binomial",
  Ntrials = c(dat$total, rep(1, nrow(pred_data))),
  data    = inla.stack.data(stack_full),
  control.predictor = list(
    A       = inla.stack.A(stack_full),
    compute = TRUE,
    link    = 1
  )
)

idx_pred <- inla.stack.index(stack_full, tag = "prediction")$data
pred <- fit_pred$summary.fitted.values[idx_pred, ]

```


# Exceedance probability: P(prevalence > threshold)

```{r}
#| label: exceedance-probability
#| message: false
#| warning: false
#| cache: false
#| include: true

exceed_prob <- vapply(
  fit_pred$marginals.fitted.values[idx_pred],
  function(m) 1 - inla.pmarginal(exceedance_threshold, m),
  numeric(1)
)

```


# WRITE PREDICTIONS BACK INTO RASTERS

```{r}
#| label: write-predictions
#| message: false
#| warning: false
#| cache: false
#| include: true

fill_raster <- function(template, idx, vals) {
  out <- template
  v <- rep(NA_real_, terra::ncell(template))
  v[idx] <- vals
  terra::values(out) <- v
  out
}

r_mean    <- fill_raster(r_mask, idx_in, pred$mean)
r_sd      <- fill_raster(r_mask, idx_in, pred$sd)
r_lower   <- fill_raster(r_mask, idx_in, pred$`0.025quant`)
r_upper   <- fill_raster(r_mask, idx_in, pred$`0.975quant`)
#r_exceed  <- fill_raster(r_mask, idx_in, exceed_prob)
r_ciwidth <- r_upper - r_lower   

pred_stack <- c(r_mean, r_sd, r_lower, r_upper, r_ciwidth)
names(pred_stack) <- c("mean", "sd", "lower95", "upper95", "ci_width")

terra::writeRaster(
  pred_stack,
  "./Data/Derived_data/vaccination_prediction_surface.tif",
  overwrite = TRUE)

```

# MAPS

```{r}
#| label: maps
#| message: false
#| warning: false
#| cache: false
#| include: true
#| 
df_mean   <- as.data.frame(r_mean,    xy = TRUE, na.rm = TRUE)
names(df_mean)[3] <- "prevalence"

df_ciw    <- as.data.frame(r_ciwidth, xy = TRUE, na.rm = TRUE)
names(df_ciw)[3] <- "ci_width"

#df_exceed <- as.data.frame(r_exceed,  xy = TRUE, na.rm = TRUE)
#names(df_exceed)[3] <- "p_exceed"

immunisation_predicted_probs <- ggplot() +
  geom_raster(
    data = df_mean, 
    aes(x, y, fill = prevalence)) +
  geom_sf(
    data = kenya, 
    fill = NA, 
    colour = "white",
     linewidth = 0.4,
     inherit.aes = FALSE) +
  scale_fill_viridis_c(
    name    = "Prevalence",
    option  = "H",
    direction = -1,
    labels  = scales::percent_format(accuracy = 1),
    limits  = c(0, max(df_mean$prevalence, na.rm = TRUE))) +
  coord_sf(crs = st_crs(kenya), expand = FALSE) +
  labs(
    subtitle = "Full vaccination") +
  theme_minimal(base_family = "Candara", base_size = 13) +
  theme(
    axis.text  = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = "gray96"),
    panel.grid = element_blank(),
    plot.subtitle = element_text(hjust = .55, vjust = -10))

immunisation_predicted_probs 

ggsave(
  "./Plots/Vaccination_predicted_prevalence.png",
  immunisation_predicted_probs,
  width = 5,
  height = 5,
  dpi = 500,
  bg = "white")

```