---
title: "Task 4: Where to Target Interventions"
subtitle: "County prioritization from Task 2's stunting evidence"
author: "[assign an owner -- see PLAN.md]"
date: today
---

<!--
Reads only ../data/processed/task2_stunting_by_county.csv and
task4_priority_counties.csv (scripts/intervention_prioritization.R,
`make intervention`) -- both aggregate, non-restricted. No KDHS microdata
access is required to render this report.
-->

```{r}
#| label: setup
#| message: false
#| warning: false
#| include: false

pacman::p_load(tidyverse, gt)
priority <- read_csv("../data/processed/task4_priority_counties.csv", show_col_types = FALSE)
national_pct <- 17.4  # data/processed/task2_stunting_summary.json -> national.stunting_prevalence_weighted_pct
threshold <- round(national_pct + sd(priority$prevalence_pct), 1)
flagged <- priority |> filter(priority)
```

## Prioritization rule

A county is flagged for priority intervention if its survey-weighted
stunting prevalence is at or above **national prevalence + 1 standard
deviation across counties** (`r national_pct`% + `r round(sd(priority$prevalence_pct), 1)`
pts = **`r threshold`%**) — meaningfully worse than typical variation, not
merely above the national average. This is a v1 rule deliberately built on
Task 2's most complete indicator (stunting, all 47 counties) rather than
waiting on wealth-quintile/immunisation/SBA data to land; see
`quarto/wealth_analysis.qmd` for how those axes will refine this ranking
once available.

## Flagged counties

```{r}
#| label: tbl-priority
#| echo: false
flagged |>
  select(priority_rank, county, prevalence_pct, ci_lower, ci_upper, gap_vs_national_pct) |>
  gt() |>
  cols_label(
    priority_rank = "Rank", county = "County", prevalence_pct = "Prevalence (%)",
    ci_lower = "95% CI lower", ci_upper = "95% CI upper", gap_vs_national_pct = "vs. national (pts)"
  ) |>
  tab_header(title = sprintf("%d counties at or above %.1f%% stunting prevalence", nrow(flagged), threshold))
```

```{r}
#| label: fig-priority
#| echo: false
#| fig-cap: "All 47 counties, ranked; flagged counties shaded."
priority |>
  ggplot(aes(x = fct_reorder(county, prevalence_pct), y = prevalence_pct, fill = priority)) +
  geom_col() +
  geom_hline(yintercept = national_pct, linetype = "dashed", color = "grey40") +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#B23A48", "FALSE" = "grey70"), guide = "none") +
  labs(x = NULL, y = "Stunting prevalence (%)") +
  theme_bw(base_size = 10)
```

## Reading the confidence intervals before acting

Some flagged counties (see `ci_width_pct` in the underlying CSV) have wide
confidence intervals — a signal that the estimate rests on fewer sampled
clusters. Two counties with the same point estimate are not equally strong
cases for intervention if one has a CI half the width of the other. Any
resourcing decision should weight this alongside the point estimate, not
read the ranking as if every row carries equal certainty.

## What would sharpen this ranking further

- Fold in `quarto/wealth_analysis.qmd`'s poorest-quintile gap once computed
  — a county that is both high-stunting and has a wide internal wealth gap
  is a stronger case than stunting alone suggests.
- Add immunisation and skilled-birth-attendance prevalence per county, once
  available, as corroborating (or contradicting) evidence — a county could
  rank moderately on stunting alone but poorly across all three indicators.
- `data/external/` is reserved for supporting context (e.g. facility density
  or distance-to-care layers) that isn't in KDHS itself, if pursued as a
  stretch goal.
