---
title: "Task 2: Wealth Analysis — Indicator Gaps by Wealth Quintile"
subtitle: "Stunting, immunisation, and skilled birth attendance across Kenya's DHS wealth index"
author: "Kevinson Mwangi & Elphas Abok"
date: today
---

<!--
Reads ../data/processed/task2_wealth_quintile.csv, produced by
scripts/wealth_quintile_analysis.R (make wealth). That script is written
and ready but has not yet been run against the real KDHS file -- this
report renders a clear "pending" state until it has, rather than failing
or (worse) showing fabricated numbers.
-->

```{r}
#| label: setup
#| message: false
#| warning: false
#| include: false

pacman::p_load(tidyverse, gt)
wealth_path <- "../data/processed/task2_wealth_quintile.csv"
has_data <- file.exists(wealth_path)
if (has_data) wealth <- read_csv(wealth_path, show_col_types = FALSE)
```

## Status

```{r}
#| echo: false
#| results: asis
if (!has_data) {
  cat(
    "**Not yet generated.** Run `make wealth` ",
    "(`scripts/wealth_quintile_analysis.R`) against the real KDHS file to ",
    "produce `data/processed/task2_wealth_quintile.csv`. The analysis code ",
    "is complete and reuses the same survey design object as ",
    "`regional_analysis.R` -- see that script for the design setup -- it ",
    "has not been executed yet. Owners: Kevinson Mwangi, Elphas Abok."
  )
}
```

This is the highest-priority remaining gap in the project: Task 2 is defined
as indicators varying by **region and wealth quintile**, and the wealth axis
currently has no coverage anywhere in the repository. Closing this gap
matters more for the project's completeness than any further regional
breakdown, since region alone only answers half of what Task 2 asks.

## Wealth-quintile estimates

```{r}
#| label: fig-wealth
#| echo: false
#| eval: !expr has_data
wealth |>
  mutate(wealth_quintile = factor(wealth_quintile, levels = c("Poorest", "Poorer", "Middle", "Richer", "Richest"))) |>
  ggplot(aes(x = wealth_quintile, y = prevalence_pct, color = indicator, group = indicator)) +
  geom_line() + geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  labs(x = NULL, y = "Prevalence (%)", color = "Indicator") +
  theme_bw(base_size = 13)
```

```{r}
#| label: tbl-wealth
#| echo: false
#| eval: !expr has_data
wealth |>
  gt(groupname_col = "indicator") |>
  cols_label(wealth_quintile = "Wealth quintile", prevalence_pct = "Prevalence (%)",
             ci_lower = "95% CI lower", ci_upper = "95% CI upper")
```

## Why this matters for Task 4

A poorest-vs-richest quintile gap, once computed, is a national-level equity
signal that complements the county-level geography in
`quarto/regional_analysis.qmd` — a county can look average on stunting while
still masking a sharp within-county wealth gap. `scripts/intervention_prioritization.R`
already checks for this file and folds a one-line wealth-gap note into the
Task 4 output if present.
